---
title: 'Econ 573: Problem Set 4'
author: "Harvey Duperier"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries}
library(ISLR2)
library(gbm)
library(glmnet)
library(randomForest)
library(MASS)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(broom)
library(knitr)
library(caret)
library(splines)
library(boot)
```
# Part I
## Question 3 Chapter 5

**3a).** The k-fold cross-validation is implemented by dividing the set of observations into k folds of approximately equal size. This first fold is treated as a validation set, and the method is fit on the remaining k-1 folds. The mean squared error is then computed on the observations in the held-out gold. This procedure is repeated k times; each time, a different group of observations is treated as a validation set. This process results in k estimates of the test error, $$MSE_1$$, $$MSE_2$$, …, $$MSE_k$$. The k-fold CV estimate is computer by averaging these values. 

**3b).** The advantages of k-fold CV relative to the validation set approach is that the validation set approach error rate may tend to overestimate the test error rate for the model fit on the entire data set. The disadvantages of the k-fold CV relative to the validation set approach is that the validation set approach is conceptually simpler than k-fold and easily implemented as it only involves partitioning the existing training data into two sets. The advantages of k-fold CV relative to LOOCV is computation. Since LOOCV is a special case of k-fold CV in which k is set to equal n, it has the potential to be computationally expensive, whereas performing 10-fold CV requires fitting the learning procedure only ten times, which could be much more feasible. The disadvantages of k-fold CV relative to LOOCV is that LOOCV has a higher variance, but lower bias, than the k-fold CVs.

## Question 5 Chapter 5
**5a).**
```{r 5a}
data("Default")
set.seed(1)
glm.fit <- glm(default ~ income + balance, data = Default, family = binomial)
summary(glm.fit)
```

**5bi).**
```{r 5bi}
train = sample(dim(Default)[1], dim(Default)[1] / 2)
```

**5bii).**
```{r 5bii}
fit.glm = glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
summary(fit.glm)
```

**5biii).**
```{r 5biii}
glm.probs = predict(fit.glm, newdata = Default[-train, ], type="response")
glm.pred=rep("No",5000)
glm.pred[glm.probs>0.5] = "Yes"
```

**5biv).**
```{r 5biv}
mean(glm.pred != Default[-train, ]$default)
```

**5c).**
```{r 5ci}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

```{r 5cii}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

```{r 5ciii}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm <- rep("No", length(probs))
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```

From the work above, we see that the validation estimate of the Test Error can be variable, depending on the difference in which observations are included in the training and validation set.

**5d).**
```{r 5d}
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance + student, data = Default, family = "binomial", subset = train)
pred.glm <- rep("No", length(probs))
probs <- predict(fit.glm, newdata = Default[-train, ], type = "response")
pred.glm[probs > 0.5] <- "Yes"
mean(pred.glm != Default[-train, ]$default)
```
Based off the results above, the student dummy variable does not lead to a reduction in the validation set estimate of the test error rate.

## Question 6 Chapter 5
**6a).**
```{r 6a}
set.seed(1)
train <- sample(dim(Default)[1], dim(Default)[1] / 2)
fit.glm <- glm(default ~ income + balance, data = Default, family = "binomial", subset = train)
summary(fit.glm)
```
The standard errors are given in the summary above from fit.glm.

**6b).**
```{r 6b}
boot.fn <- function(data, index) {
    fit <- glm(default ~ income + balance, data = data, family = "binomial", subset = index)
    return (coef(fit))}
```

**6c).**
```{r 6c}
boot(Default, boot.fn, 1000)
```
The bootstrap of the estimates of the standard errors are listed above in the summary.

**6d).**
Overall, the estimate standard errors obtained by the two differing methods are pretty close and there wouldn't be a huge difference in using one method over the other, but I personally would chose to use the bootstrap in this case.


# Part II
## Question 3 Chapter 7
```{r q3}
x = -2:2
y = 1 + x + -2 * (x-1)^2 * I(x>1)
plot(x, y)
```

## Question 9 Chapter 7
**9a).**
```{r 9a}
data("Boston")
set.seed(1)
theme_set(theme_tufte(base_size = 20) + theme(legend.position = 'top'))
data('Boston')

model <- lm(nox ~ poly(dis, 3), data = Boston)
tidy(model) %>%
  kable(digits = 3)
Boston %>%
  mutate(pred = predict(model, Boston)) %>%
  ggplot() +
  geom_point(aes(dis, nox, col = '1')) +
  geom_line(aes(dis, pred, col = '2'), size = 1) +
  scale_color_manual(name = 'Value Type/Color:',
                     labels = c('Observed', 'Predicted'),
                     values = c('#7BAFD4', '#8B0000'))
```
Each power of the dis coefficient is found to be statistically significant according to the model. The plot also seems to describe the data without overfitting.

**9b).**
```{r 9b}
errors <- list()
models <- list()
pred_df <- data_frame(V1 = 1:506)
for (i in 1:9) {
  models[[i]] <- lm(nox ~ poly(dis, i), data = Boston)
  preds <- predict(models[[i]])
  pred_df[[i]] <- preds
  errors[[i]] <- sqrt(mean((Boston$nox - preds)^2))
}
errors <- unlist(errors)
names(pred_df) <- paste('Level', 1:9)
data_frame(RMSE = errors) %>%
  mutate(Poly = row_number()) %>%
  ggplot(aes(Poly, RMSE, fill = Poly == which.min(errors))) +
  geom_col() + 
  guides(fill = "none") +
  scale_x_continuous(breaks = 1:9) +
  coord_cartesian(ylim = c(min(errors), max(errors))) +
  labs(x = 'Degree Rate')
Boston %>%
  cbind(pred_df) %>%
  gather(Polynomial, prediction, -(1:14)) %>%
  mutate(Polynomial = factor(Polynomial, 
                             levels = unique(as.character(Polynomial)))) %>%
  ggplot() + 
  ggtitle('Predicted Values for Each Level of Polynomial') +
  geom_point(aes(dis, nox, col = '1')) + 
  geom_line(aes(dis, prediction, col = '2'), size = 1) +
  scale_color_manual(name = 'Value Type',
                     labels = c('Observed', 'Predicted'),
                     values = c('#7BAFD4', '#8B0000')) +
  facet_wrap(~ Polynomial, nrow = 2)
```
From this we see that the model with the highest polynomial degree has the lowest RSS when fitted and tested on the same data.

**9c).**
```{r 9c}
errors <- list()
folds <- sample(1:10, 506, replace = TRUE)
errors <- matrix(NA, 10, 9)
for (k in 1:10) {
  for (i in 1:9) {
    model <- lm(nox ~ poly(dis, i), data = Boston[folds != k,])
    pred <- predict(model, Boston[folds == k,])
    errors[k, i] <- sqrt(mean((Boston$nox[folds == k] - pred)^2))
  }
}
errors <- apply(errors, 2, mean)
data_frame(RMSE = errors) %>%
  mutate(Poly = row_number()) %>%
  ggplot(aes(Poly, RMSE, fill = Poly == which.min(errors))) +
  geom_col() + theme_tufte() + guides(fill = FALSE) +
  scale_x_continuous(breaks = 1:9) +
  coord_cartesian(ylim = range(errors))
```
The model with a polynomial degree of 4 is chosen when tested on out-of-sample data. This is the highest polynomial degree that does not show signs of overfitting like 5 through 9 do.

**9d).**
```{r 9d}
model <- lm(nox ~ bs(dis, df = 4), data = Boston)
kable(tidy(model), digits = 3)
Boston %>%
  mutate(pred = predict(model)) %>%
  ggplot() +
  geom_point(aes(dis, nox, col = '1')) + 
  geom_line(aes(dis, pred, col = '2'), size = 1.5) +
  scale_color_manual(name = 'Value Type',
                     labels = c('Observed', 'Predicted'),
                     values = c('#7BAFD4', '#8B0000')) +
  theme_tufte(base_size = 13)
```
This model finds all the different bases to be statistically significant and the pred line seems to fit data well without overfitting.

**9e).**
```{r 9e}
errors <- list()
models <- list()
pred_df <- data_frame(V1 = 1:506)
for (i in 1:9) {
  models[[i]] <- lm(nox ~ bs(dis, df = i), data = Boston)
  preds <- predict(models[[i]])
  pred_df[[i]] <- preds
  errors[[i]] <- sqrt(mean((Boston$nox - preds)^2))
}

names(pred_df) <- paste(1:9, 'Degrees of Freedom')
data_frame(RMSE = unlist(errors)) %>%
  mutate(df = row_number()) %>%
  ggplot(aes(df, RMSE, fill = df == which.min(errors))) +
  geom_col() + guides(fill = FALSE) + theme_tufte() +
  scale_x_continuous(breaks = 1:9) +
  coord_cartesian(ylim = range(errors))
Boston %>%
  cbind(pred_df) %>%
  gather(df, prediction, -(1:14)) %>%
  mutate(df = factor(df, levels = unique(as.character(df)))) %>%
  ggplot() + ggtitle('Predicted Values for Each Level of Polynomial') +
  geom_point(aes(dis, nox, col = '1')) + 
  geom_line(aes(dis, prediction, col = '2'), size = 1.5) +
  scale_color_manual(name = 'Value Type',
                     labels = c('Observed', 'Predicted'),
                     values = c('#7BAFD4', '#8B0000')) +
  facet_wrap(~ df, nrow = 3)
```
When trained and tested on the same data, the higher complexity models are deemed the best as shown on the plots above.

**9f).**
```{r 9f}
folds <- sample(1:10, size = 506, replace = TRUE)
errors <- matrix(NA, 10, 9)
models <- list()
for (k in 1:10) {
  for (i in 1:9) {
    models[[i]] <- lm(nox ~ bs(nox, df = i), data = Boston[folds != k,])
    pred <- predict(models[[i]], Boston[folds == k,])
    errors[k, i] <- sqrt(mean((Boston$nox[folds == k] - pred)^2))
  }
}
errors <- apply(errors, 2, mean)
data_frame(RMSE = errors) %>%
  mutate(df = row_number()) %>%
  ggplot(aes(df, RMSE, fill = df == which.min(errors))) +
  geom_col() + theme_tufte() + guides(fill = FALSE) +
  scale_x_continuous(breaks = 1:9) +
  coord_cartesian(ylim = range(errors))
```
These were validated on out-of-sample data. Due to this, a simpler model is chosen, with df of 4. This is similar to polynomial validation, in that this is the most complex model that does not begin to show signs of overfitting like 5 through 9 do.


# Part III
## Question 1 Chapter 8
```{r 1C8}
par(xpd=NA)
plot(NA, NA, type="n", xlim=c(0,100), ylim=c(0,100), xlab="X", ylab="Y")
# t1: x = 40; (40, 0) (40, 100)
lines(x=c(40,40),y=c(0,100))
text(x=40, y=108, labels=c("t1"), col="red")
# t2: y = 75; (0, 75) (40, 75)
lines(x=c(0,40), y=c(75,75))
text(x=-8, y=75, labels=c("t2"), col="red")
# t3: x = 75; (75,0) (75, 100)
lines(x=c(75,75),y=c(0,100))
text(x=75, y=108, labels=c("t3"), col="red")
# t4: x = 20; (20,0) (20, 75)
lines(x=c(20,20),y=c(0,75))
text(x=20, y=80, labels=c("t4"), col="red")
# t5: y=25; (75,25) (100,25)
lines(x=c(75,100),y=c(25,25))
text(x=70, y=25, labels=c("t5"), col="red")
text(x=(40+75)/2, y=50, labels=c("R1"))
text(x=20, y=(100+75)/2, labels=c("R2"))
text(x=(75+100)/2, y=(100+25)/2, labels=c("R3"))
text(x=(75+100)/2, y=25/2, labels=c("R4"))
text(x=30, y=75/2, labels=c("R5"))
text(x=10, y=75/2, labels=c("R6"))


       "[  X<40 ] 
        |       |
    [Y<75]    [X<75]
    |   |     |    |
 [X<20] R2    R1   [Y<25]
 |    |            |    |
 R6   R5           R4   R3"
        
```
      
## Question 5 Chapter 8

In the majority approach, the final classification is red since the number of red predictions is greater than the number of green predictions. In the second approach, the final classification is green since the average of the probabilities is less than the threshold of 50%.

## Question 10 Chapter 8
**10a).**
```{r 10a}
data("Hitters")
Hitters = na.omit(Hitters)
Hitters$Salary = log(Hitters$Salary)
```

**10b).**
```{r 10b}
train = 1:200
hitters.train = Hitters[train,]
hitters.test = Hitters[-train,]
```

**10c).**
```{r 10c}
set.seed(1)
pows = seq(-10, -0.2, by = 0.1)
lambdas = 10^pows
train.err = rep(NA, length(lambdas))
for (i in 1:length(lambdas)) {
  boost.hitters = gbm(Salary ~ ., data = hitters.train, distribution = "gaussian", n.trees = 1000, shrinkage = lambdas[i])
  pred.train = predict(boost.hitters, hitters.train, n.trees = 1000)
  train.err[i] = mean((pred.train - hitters.train$Salary)^2)
}
plot(lambdas, train.err, type = "b", xlab = "Shrinkage-values", ylab = "Training-MSE")
```

**10d).**
```{r 10d}
set.seed(1)
test.err <- rep(NA, length(lambdas))
for (i in 1:length(lambdas)) {
  boost.hitters = gbm(Salary ~ ., data = hitters.train, distribution = "gaussian", n.trees = 1000, shrinkage = lambdas[i])
  yhat = predict(boost.hitters, hitters.test, n.trees = 1000)
  test.err[i] = mean((yhat - hitters.test$Salary)^2)
}
plot(lambdas, test.err, type = "b", xlab = "Shrinkage-values", ylab = "Test-MSE")
min(test.err)
lambdas[which.min(test.err)]
```

**10e).**
```{r 10e}
fitFirst = lm(Salary ~ ., data = hitters.train)
predFirst = predict(fitFirst, hitters.test)
mean((predFirst - hitters.test$Salary)^2)
x = model.matrix(Salary ~ ., data = hitters.train)
x.test = model.matrix(Salary ~ ., data = hitters.test)
y = hitters.train$Salary
fitSnd = glmnet(x, y, alpha = 0)
predSnd = predict(fitSnd, s = 0.01, newx = x.test)
mean((predSnd - hitters.test$Salary)^2)
```
The test MSE for boosting is lower than for linear and ridge regression shown by the data above.

**10f).**
```{r 10f}
boost.hitters <- gbm(Salary ~ ., data = hitters.train, distribution = "gaussian", n.trees = 1000, shrinkage = lambdas[which.min(test.err)])
summary(boost.hitters)

```
The variable 'CAtBat' is the most important predictor in the boosted model.

**10g).**
```{r 10g}
set.seed(1)
bag.hitters <- randomForest(Salary ~ ., data = hitters.train, mtry = 19, ntree = 500)
yhat.bag <- predict(bag.hitters, newdata = hitters.test)
mean((yhat.bag - hitters.test$Salary)^2)
```
The test set MSE for this approach is 0.2299324 which is slightly lower than the test MSE for boosting.








